---
title: "MATH 588"
subtitle: "HW6"
author: "Md Ismail Hossain"
date: "3/27/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

\newpage

# Question 1 

## a

```{r}
library(Sleuth3)
mem = case1601



names(mem) = casefold(names(mem))
# Group means and SEs as 2 by 5 matrices
means = as.matrix(aggregate(mem[,3:7], list(mem$treatment), mean)[,-1])
sds = as.matrix(aggregate(mem[,3:7], list(mem$treatment), sd)[,-1])
ns = as.matrix(aggregate(mem[,3:7], list(mem$treatment), length)[,-1])
SEs = sds/sqrt(ns-1)
multiplier = matrix(qt(0.975, ns-1), ncol=5, nrow=2)
upper = means + multiplier*SEs
lower = means - multiplier*SEs
# Start with an empty plot that will hold everything
lim = c(min(lower), max(upper))

{plot(c(2,16), lim, type="n", xlab="Training Week Prior to Treatment",
ylab="Percentage Correct",
main="Monkey Memory Study")
# Means and CIs by treatment
x = c(2,4,8,12,16)
delta = 0.1
epsilon = 0.075
for (i in 1:2) {
  xa = x+c(-1,1)[i]*delta
  lines(xa, means[i,], col=i, lty=i, pch=i, type="b")
  segments(xa, lower[i,], xa, upper[i,], col=i)
  segments(rep(xa-epsilon,2), c(lower[i,],upper[i,]),
  rep(xa+epsilon,2), c(lower[i,],upper[i,]), col=i)
}
# Legend
legend(9, 90, c("Treated","Untreated"), lty=2:1, col=2:1, pch=2:1)
# Optional "CI" box:
text(c(14,14), c(52.5,51), c("95%","C.I."), adj=0)
segments(c(15.25,15.4,15.25), c(50.5,50.5,52.5), c(15.55,15.4,15.55), c(50.5,52.5,52.5))
points(15.4,51.5)
rect(13.8, 49.5, 15.8, 53.6)}

```




## b

```{r}
CTRL=mem$treatment=="Control"
round(diag(cov(mem[CTRL,3:7])),3)
round(diag(cov(mem[!CTRL,3:7])),3)

## OR just sds^2 from above
apply(sds^2, 2, function(x) max(x)/min(x))

round(abs(cor(mem[!CTRL,3:7]) - cor(mem[CTRL,3:7])) ,3)
```
Week 12 has a highest variance between controls and treated group which is $5.39$ times higher for controls.The highest correlation values  between pairs week 4 and 16 and it is $0.844$. (between control and treated group)

## c

```{r}
mem1 = aov(as.matrix(mem[,3:7])~treatment, data=mem)
anova(mem1, test="Hotelling")
```
 

```{r}
anova(aov(cbind(week2,week4,week8,week12,week16)~treatment, data=mem))
```


The p-value is less than 0.05. So, the null hypothesis (means are equal) can be rejected at 5\% level of significance.

## d

```{r}
source("http://www.stat.cmu.edu/~hseltman/files/qqn.R")
qqn(resid(mem1), main="Monkey Memory Quantile Normal Plot")
```
The plot suggest, the normality assumption does not violated in this case. 

# Question 2 

## a

```{r}
nn = ex1605
names(nn) = casefold(names(nn))
{par(mfrow=c(2,2), oma=c(0,0,1.5,0))
  
with(nn, plot(fmed, age2iq, xlab="Foster Mother's Education",
ylab="Child's IQ at 2 years"))
with(nn, plot(tmiq, age2iq, xlab="Birth Mother's IQ",
ylab="Child's IQ at 2 years"))
with(nn, plot(fmed, age13iq, xlab="Foster Mother's Education",
ylab="Child's IQ at 13 years"))

with(nn, plot(tmiq, age13iq, xlab="Birth Mother's IQ",
ylab="Child's IQ at 13 years"))
mtext("Nature / Nurture EDA", outer=T, cex=1.3)}
```

## b

```{r}
library(car)
idata = data.frame(time = ordered(1:4, labels=paste("age",c(2,4,8,13),sep="")))
rslt = Manova(lm(cbind(age2iq,age4iq,age8iq,age13iq)~fmed+tmiq,nn),
idata=idata, idesign=~time)
print(rslt)
```

```{r}
anova(aov(cbind(age2iq,age4iq,age8iq,age13iq)~fmed+tmiq,nn))
```



In the model we observed that the p-value for variable time "time" is less than 0.05. As a result we can interpret that at 5\% level of significance we can reject the null hypothesis (average IQ similar over the time) considering all other factors are constant. 


## c

The null hypothesis that the "IQs do not change with the birth mothers
IQ" can be rejected at 5\% level of significance because the p-value is less than 0.05. We observed a positive correlation between mother's age and age 13 group, but other plots are not very clear to make a statement. 

# Question 3


```{r}
library(readr)
fakebeet <- read_csv("E:/NMT MS/Spring 22/MATH 588/Home_Work/Spring-2022---MATH-588-01-Advanced-Data-Analysis/HW6/fakebeet.csv")
with(fakebeet, t.test(P[species=="one"], P[species=="two"]))

with(fakebeet, t.test(Q[species=="one"], Q[species=="two"]))

anova(aov(cbind(P,Q)~species, data=fakebeet), test="Hotelling")
```

The individual t-test are presenting a p-value of greater than 0.05, which interpreting that the null hypothesis can not be rejected. But the multivariate T square p-value is less than 0.05. So, it's suggest that we can reject the null hypothesis that individual species are equal. That's why multivariate t-test is important to consider. 


# Question 4

```{r}
library(readr)
swiss1 <- read_table2("E:/NMT MS/Spring 22/MATH 588/Home_Work/Spring-2022---MATH-588-01-Advanced-Data-Analysis/HW6/swiss1.txt")


genuine <-  swiss1[1:100,2:7]
genuine$Type <- rep("Genuine",100)
counterfeit  <-  swiss1[101:200,2:7]
counterfeit$Type <- rep("Counterfeit",100)
dat <- rbind(genuine,counterfeit)
```


```{r}
mod <- manova(cbind(Length,Left,Right,Bottom,Top,Diagonal) ~ Type, data=dat)
```


```{r}
summary(mod)
```


We can reject the null hypothesis at 5\% level of significance because the p-value is less than $0.05$. 

```{r}

ci_ellip_plt <- function(var){
  
  library(car)
  #var = "Length"
  x = dat[dat$Type == "Genuine",var]
  y = dat[dat$Type == "Counterfeit",var]
  cid = data.frame(x,y)
  
  colnames(cid) <- c(paste(var,"Genuine"),paste(var,"Counterfeit"))
  
  
  n=100
  p=2
  
  Sd=cov(cid)
  dbar=c(colMeans(cid))
  
  # Plot the 95% confidence ellipse for the vector of mean differences
  library(ellipse)
  plot(ellipse(Sd,centre=dbar,t=sqrt(((n-1)*p/(n*(n-p)))*qf(0.95,p,n-p))),type="l",main = paste("95% CI Ellipsoid for",var))
  
}
par(mfrow=c(3,2))
ci_ellip_plt(var = "Length")
ci_ellip_plt(var = "Left")
ci_ellip_plt(var = "Right")
ci_ellip_plt(var = "Bottom")
ci_ellip_plt(var = "Top")
ci_ellip_plt(var = "Diagonal")
```

