---
title: "MATH 588"
subtitle: "HW3"
author: "Md Ismail Hossain"
date: "2/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

\newpage

# Question: 4.29


\textbf{Null Hypothesis:} Logging has no effect on the distribution of percentage seedling lose lose between 2004 and 2005. \newline
\textbf{Alternative Hypothesis:} Logging has statistically significant effect on the distribution of percentage seedling lose between 2004 and 2005.


```{r}
library(Sleuth3)
head(ex0429)

# Median Percent Lost by Action
tapply(ex0429$PercentLost,ex0429$Action,median,na.rm=T)

# a
# Rank sum test
wilcox.test(PercentLost~Action,conf.int = TRUE,data = ex0429)




```

The p-value $(0.01154)$ is less than 0.05. So, at 5\% level of significance we can
reject the null hypothesis and conclude than logging has significant impact
on the distribution of percentage of seedlings between 2004 and 2005.


The 95\% confidence interval on the difference in median is $(10.8,65.1)$

```{r}
#b
t.test(PercentLost~Action,var.equal=TRUE,data = ex0429)
```

Comparing the rank-sum test results and independent sample t-test results, both giving
us the same decision of rejecting the null hypothesis. But the 95\% CI interval on the difference
in the medians using rank-sum little smaller than the t-tools CI with same level
of significance.

# Question 5.15

```{r}


# a
sample_var = var(case0502$Percent)

# b
total_ss = sample_var*45
total_ss
# c

sp = ((9-1)*5.03^2+(5-1)*11.94^2+(6-1)*6.58^2+(9-1)*4.59^2+(2-1)*3.818^2+(6-1)*9.010^2+(9-1)*5.969^2)/((9-1)+(5-1)+(6-1)+(9-1)+(2-1)+(6-1)+(9-1))
residual_ss = sp*(46-7)
residual_ss

ms_resid = residual_ss/(46-7)
ms_resid

# SS Between groups
ss_btwn = total_ss-residual_ss
ss_btwn

ms_btwn = ss_btwn/6
ms_btwn


# F-statistic
f_statistic = ms_btwn/ms_resid
f_statistic

```






\
\
\
\
\
\
\
\
\
```{r}
#d
pf(f_statistic,6,39, lower.tail = FALSE)
```
So, we are getting exactly similar results like in Disply 5.10


# Question: 5.21


\textbf{Null Hypothesis:} All population variance are equal. \newline
\textbf{Alternative Hypothesis:} Populaiton variance are unequal.

```{r}
#install.packages("car")
library(car)
# Levene's test with one independent variable
leveneTest(Percent ~ Judge, data = case0502)
```

At 5\% level of significance we can not reject the null hypothesis because p-value $(0.2969)$ is higher than $0.05$ and conclude that population variance are equal.


# Question: 6.23

```{r}

head(ex0623)

# install.packages("ggplot2")
library(ggplot2)

ggplot(ex0623, aes(x = WtLoss24)) +
  geom_histogram(fill = "white", colour = "black") +
  facet_grid(Group ~ .)








```

The histogram suggest that data does not suffer severally from non normality. So, we can use
one way anova to do the comparison among groups. 

```{r}
# Compute the analysis of variance
anova_oneway <- aov(WtLoss24 ~ Group, data = ex0623)
# Summary of the analysis
summary(anova_oneway)

```

The p-value $0.0409$ is less than $0.05$, so at 5\% level of significance we can conclude that
there is statistically significant difference in terms of mean weight loss between  at least one pair of the diet plan groups. 

```{r}
# Multiple test:
TukeyHSD(anova_oneway)
```
Turkeys HSD test indicates that there is statistically significant difference between  Low-Fat Vs Low-Carbohydrate diet groups.

 
```{r}
aggregate(WtLoss24 ~ Group, data = ex0623, FUN = mean)
```

And among all the diet plans, Low-Fat diet have the lowest mean weight loss. So, if losing weight is the goal for an individual then Low-Fat diet would be the winner in that case. 

## Checking ANOVA assumptions

```{r}
# Normality
plot(anova_oneway, 2)
```

The normality assumption does not looks violated. 

```{r}
# Equal variance
leveneTest(WtLoss24 ~ Group, data = ex0623)

```

This test of homogeneity suggest that, at 5\% level of significance we can conclude that the 
groups do not have equal variance. Equality of variance is one of the assumption settled before
carry out the anova. The alternative is to use non parametric approach to test the hypothesis or we could allow this inequality of the variance among groups, because in practical situation equality among groups does not found in most case. \


## Non-parametric approach:

We are getting exactly similar types of results like ANOVA even after using non-parametric approach.
```{r}
kruskal.test(WtLoss24 ~ Group, data = ex0623)
```
```{r}
pairwise.wilcox.test(ex0623$WtLoss24,ex0623$Group,
                 p.adjust.method = "BH")
```


