---
title: "MATH 588"
subtitle: "HW10"
author: "Md Ismail Hossain"
date: "4/27/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

\newpage

# Question 1 


```{r}
library(Sleuth3)
head(ex2115)
attach(ex2115)
```

## a

```{r}
ex2115$prop <- PercentInFavor/100
binResponse <- cbind(InFavor, NotInFavor)
glm1_1 <- glm (binResponse ~ as.factor(Level), data = ex2115, family = binomial)
summary (glm1_1)
```

The p-value for Level is not significant at 5\% level of significance. So , we can conclude that proportion of favorable response to the target question does not depend on the wording (LEVEL) of the question.

## b

```{r}
glm1_2 <- glm (binResponse ~ as.factor(Context), data = ex2115, family = binomial)
summary (glm1_2)

```

Here, the p-value for both the intercept and factor variable are statistically significant because the p-value is less than 0.05. The odds of beta 1 is 0.519. So, the odds of the context question preceding the target question about U.S. aid to the Nicaraguan Contra rebels of Cuba is almost 2 times than the Vietnam.


## c

```{r}
ind_mode <- ifelse (Mode == "scattered", 1, 0)
ind_context <- ifelse (Context == "Vietnam", 1, 0)
glm1_3 <- glm (binResponse ~ ind_context + ind_mode, data = ex2115, family = binomial)
summary (glm1_3)
```

Here, the two-sided p-value for the $\beta_1$ and $\beta_2$ are $6.75e-0$ and $0.00388$ respectively which are significant at
5\% level of significance. SO, we can conclude that the proportion depend on the context question to different
extents according to whether the target and context questions are scattered.


# 

```{r}
attach (ex2116)
prop_tumor <- Tumor / Total
odds_tumor <- prop_tumor / (1 - prop_tumor)
odds_tumor
```

# Exercise 21.16

```{r}
head (ex2116)
attach (ex2116)
prop_tumor <- Tumor / Total
odds_tumor <- prop_tumor / (1 - prop_tumor)
odds_tumor
```


```{r}
require(lattice)
xyplot (Dose ~ odds_tumor, data = ex2116)
```


```{r}
binResponse <- cbind (Tumor, Total-Tumor)
glm2_1 <- glm (binResponse ~ Dose, data=ex2116, family = binomial)
summary (glm2_1)
```



```{r}
ex2116$cum_rel_tumor <- cumsum(Tumor)/sum(Tumor)
ex2116$cum_rel_tumor
```


Here, the p-value is less than 0.05. So, we can conclude that there are strong relationship between dose of A flatoxicol and odds of liver tumor. From the above output we can see that at dose 0.10, 50\% of the fish will get liver tumors.

# Exercise 22.15

```{r}
ex2215 = case2201
attach(ex2215)

```

## a


```{r}
lm3_1 <- lm (sqrt(Matings) ~ Age)
summary (lm3_1)
```
The fitted simple linear regression model for the square root of the matings is given by,

$$ \hat y_i = -0.81220 + 0.06320(Age)$$

with $SE(\beta_0) = 0.56867$ and $SE(\beta_1) = 0.01561$

## b

```{r}
lm3_2 <- lm (log(Matings+1) ~ Age)
summary (lm3_2)
```

The fitted simple linear regression model for the logarithamic transformation of the matings(after adding 1)
is given by,

$$ log(\hat y_i) = -0.69893 + 0.05093(Age)$$
with $SE(\beta_0) = 0.45861$ and $SE(\beta_1) = 0.01259$


## c

```{r}
glm3_3 <- glm (Matings ~ Age, family = poisson)
summary (glm3_3)
```

With a count response matings and a explanatory variable age, the Poisson log-linear regression model is
given by,

$$ log(\hat \mu) = -1.58201 + 0.06869(Age)$$

with $SE(\beta_0) = 0.54462$ and $SE(\beta_1) = 0.01375$

## d

```{r}
par(mfrow = c(1, 2))
plot (lm3_1, 1)
plot (lm3_1, 2)
```

The residual plot will show no fitted pattern. That is, the red line should be approximately horizontal
at zero. But in that case of model fitted in (a), the red line is not horizontal at 0. In case of the normal
probability plot of residuals should approximately follow a straight line which indicate the model is adequate
but in our case for the model fitted in (a) doesn’t show the straight line rather showing some outliers at case
no 32, and 36. So from both the plot we can assume that the fitted model in (a) exhibit obvious inadequacy.

```{r}
par(mfrow = c(1, 2))
plot (lm3_2, 1)
plot (lm3_2, 2)
```

The fitted plot seems perfect to me.

# Exercise 22.22

```{r}
attach (ex2222)
glm5_1 <- glm (Count ~ Book + Word, family = poisson)
summary (glm5_1)
```
For this model, the p-value for the coefficients are less than the critical value (0.05) at 5\% level of significance which
indicates that Jane Austen was consistent in the three books in her relative uses of these words.

```{r}
ind_sandII <- ifelse (Book == "SanditonII", 1, 0)
glm5_2 <- glm (Count ~ ind_sandII + Word, family = poisson)
summary (glm5_2)
```
The emulator do a great job in terms of matching the relative rates of occurrence of these six words since the two-sided p-value is smaller than the critical value (0.05) at 5\% level of significance.


```{r}
ind_sandI <- ifelse (Book == "SanditonI", 1, 0)
glm5_3 <- glm (Count ~ ind_sandI + ind_sandII + Word, family = poisson)
summary (glm5_3)
```

From the above output we can conclude that, the emulator match the relative rates that Austen used the
words in the first part of Sanditon since the two-sided p-value is smaller than the critical value (0.05) at 5\%
level of significance.

