---
title: "MATH 588"
subtitle: "HW9"
author: "Md Ismail Hossain"
date: "4/23/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,message=FALSE)
```

\newpage

# Question 1 


```{r}
library(Sleuth3)
head(ex2115)
attach(ex2115)
```

## a

```{r}
ex2115$prop <- PercentInFavor/100
binResponse <- cbind(InFavor, NotInFavor)
glm1_1 <- glm (binResponse ~ as.factor(Level), data = ex2115, family = binomial)
summary (glm1_1)
```

The p-value for Level is not significant at 5\% level of significance. So , we can conclude that proportion of favorable response to the target question does not depend on the wording (LEVEL) of the question.

## b

```{r}
glm1_2 <- glm (binResponse ~ as.factor(Context), data = ex2115, family = binomial)
summary (glm1_2)

```

Here, the p-value for both the intercept and factor variable are statistically significant because the p-value is less than 0.05. The odds of beta 1 is 0.519. So, the odds of the context question preceding the target question about U.S. aid to the Nicaraguan Contra rebels of Cuba is almost 2 times than the Vietnam.


## c

```{r}
ind_mode <- ifelse (Mode == "scattered", 1, 0)
ind_context <- ifelse (Context == "Vietnam", 1, 0)
glm1_3 <- glm (binResponse ~ ind_context + ind_mode, data = ex2115, family = binomial)
summary (glm1_3)
```

Here, the two-sided p-value for the \β1 and \β2 are $6.75e-0$ and $0.00388$ respectively which are significant at
5\% level of significance. SO, we can conclude that the proportion depend on the context question to different
extents according to whether the target and context questions are scattered.


# 

```{r}
attach (ex2116)
prop_tumor <- Tumor / Total
odds_tumor <- prop_tumor / (1 - prop_tumor)
odds_tumor
```


